\section{A Framework built on pandas}
\label{sec:framework}

There are several motivations for building an analysis framework around the \code{pandas DataFrame}. 
The guiding principle underlying the design of this framework is to enable exploratory investigation.
This requires both intelligent memory management for handling out-of-core datasets, and a robust indexing system to ensure that particle properties do not become misaligned in memory.
Using  the \code{pandas DataFrame} as the primary data container rather than separate \code{numpy} arrays makes it much easier to keep different particle properties indexed correctly while still affording the flexibility to load and remove data from memory at will.
In addition, \code{pandas} itself is a thoroughly documented, open-source, BSD-licensed library providing high-performance, easy-to-use data structures and analysis tools, and has a strong community of developers working to improve it.  
More broadly, as \code{pandas} is becoming the de-facto standard for data analysis in python, doing so simplifies interoperability with the rest of the tools in the ecosystem.

\code{Gadfly} is designed for use with simulation data stored in the HDF5 format \citep{hdf5}.  
While we otherwise aim to keep \code{gadfly} as general as possible, some assumptions about the storage format are necessary.
Each particle type is expected to be contained in a different HDF5 group, labeled \code{PartType0, PartType1}, etc; a \code{Header} group is also expected, containing metadata for the simulation snapshot as HDF5 attributes. 
All particles are expected to have the following fields defined: particle IDs, masses, coordinates, and velocities.  
SPH particles are additionally expected to have a smoothing length, density, and internal energy.  Additional fields not included in the original \textsc{gadget} specification, such as chemical abundances, are also supported.

Here, we provide an overview of the design and capabilities of the \code{gadfly} framework, including the \code{Simulation}, \code{Snapshot}, and \code{PartType DataFrame} objects at the core of \code{gadfly} (Section \ref{sec:hierarchy}), the usage of which is demonstrated in Figure \ref{fig:usage_example}.
Our approach to file access and intelligent memory management (Section \ref{sec:fileIO}), our handling of unit conversions (Section \ref{sec:units}) and coordinate transformations (Section \ref{sec:coords}), and the included utilities for parallel batch processing (Section \ref{sec:parallel}).
